---
title: "Summary Visualizations"
output: 
  html_document:
    toc: false
    toc_float: false
    theme: flatly
---

```{r, echo = FALSE, warning= FALSE, message=FALSE}
library(tidyverse)
library(ggridges)
library(plotly)
```

## Average Temperatures by Region

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
library(rvest)

ave_temp = read_csv("https://www.ncdc.noaa.gov/cag/statewide/mapping/110-tavg.csv", skip = 3) %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("year", "month"), sep = 4) %>%
  mutate(
    year = as.numeric(year), 
    state_code = setNames(state.abb, state.name)[location]) %>% 
  filter(year >= 1953 & year <= 2018) %>% 
  group_by(location, year) %>% 
  mutate(
    mean_temp = mean(value)) %>% 
  dplyr::select(state_code, location, year, mean_temp) %>% 
  distinct()

precip = read_csv("https://www.ncdc.noaa.gov/cag/statewide/mapping/110-pcp.csv", skip = 3) %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("year", "month"), sep = 4) %>%
 mutate(
    year = as.numeric(year), 
    state_code = setNames(state.abb, state.name)[location]) %>% 
  filter(year >= 1953 & year <= 2018) %>% 
  group_by(location, year) %>% 
  mutate(
    total_precip = sum(value)) %>% 
  dplyr::select(state_code, location, year, total_precip) %>% 
  distinct()

tmax = read_csv("https://www.ncdc.noaa.gov/cag/statewide/mapping/110-tmax.csv", skip = 3) %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("year", "month"), sep = 4) %>%
 mutate(
    year = as.numeric(year), 
    state_code = setNames(state.abb, state.name)[location]) %>% 
  filter(year >= 1953 & year <= 2018) %>% 
  group_by(location, year) %>% 
  mutate(
    max_temp = max(value)) %>% 
  dplyr::select(state_code, location, year, max_temp) %>% 
  distinct()

tmin = read_csv("https://www.ncdc.noaa.gov/cag/statewide/mapping/110-tmin.csv", skip = 3)  %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("year", "month"), sep = 4) %>%
 mutate(
    year = as.numeric(year), 
    state_code = setNames(state.abb, state.name)[location]) %>% 
  filter(year >= 1953 & year <= 2018) %>% 
  group_by(location, year) %>% 
  mutate(
    min_temp = min(value)) %>% 
  dplyr::select(state_code, location, year, min_temp) %>% 
  distinct()

disasters = read_csv("./data/DisasterDeclarationsSummaries2.csv") %>%
  janitor::clean_names() %>%
  mutate(disaster = factor(incident_type)) %>%
  rename(
     year = fy_declared) %>%
  filter(year >= 1953 & year <= 2018) %>%
  count(state, year, disaster) %>%
  ungroup()

final_data =
  inner_join(ave_temp, precip, by = c("location" = "location", "year" = "year", "state_code" = "state_code")) %>%
  inner_join(tmax, by = c("location" = "location", "year" = "year", "state_code" = "state_code")) %>%
  inner_join(tmin, by = c("location" = "location", "year" = "year", "state_code" = "state_code")) %>%
  full_join(disasters, by = c("year" = "year", "state_code" = "state"))

final_data = final_data %>% 
  mutate(n = replace_na(n, 0)) %>%
  group_by(year, location) %>%
  mutate(
    n_state = sum(n)
  ) %>%
  group_by(year) %>%
  mutate(
    n_total = sum(n)
  ) %>%
  group_by(year, disaster) %>%
  mutate(
    n_type = sum(n)
  ) %>%
  rename("n_disaster_state" = "n") %>%
  mutate(
  region = case_when(
    location %in% c(
                    "Alabama","Arkansas","Delaware","Florida","Georgia"
                    ,"Kentucky","Louisiana","Maryland","Mississippi",
                    "Oklahoma","North Carolina","South Carolina",
                    "Tennessee","Texas","Virginia","West Virginia") ~ "southeast",
   location %in% c("Connecticut","Maine","New Hampshire","Massachusetts",
                   "New Jersey","New York","Pennsylvania","Rhode Island","Vermont") ~ 'northeast',
  location %in% c("Alaska",
                "Arizona","California","Colorado","Hawaii","Idaho",
                "Montana","Nevada","New Mexico","Oregon","Utah","Washington","Wyoming")
       ~  'west',
  location %in% c("Illinois",
                "Indiana","Iowa","Kansas","Michigan","Missouri",
                "Minnesota","Nebraska","North Dakota","Ohio","South Dakota", "Wisconsin") ~ 'midwest'))
 
regional_data = final_data %>%
  group_by(year, region) %>%
  mutate(
    ave_temp = mean(mean_temp),
    sum_precip = sum(total_precip),
    count_region = sum(n_disaster_state)
  ) %>%
  select(region, year, ave_temp, sum_precip, count_region) %>%
  filter(region != "NA") %>%
  distinct() %>%
  ungroup()

regional_disaster_count = final_data %>%
  group_by(disaster, year, region) %>%
  mutate(
    ave_temp = mean(mean_temp),
    sum_precip = sum(total_precip),
    count_region = sum(n_disaster_state)
  ) %>%
  select(region, year, ave_temp, sum_precip, count_region) %>%
  filter(region != "NA" & disaster != "NA") %>%
  distinct() %>%
  ungroup()
```



```{r, scatterplot, echo = FALSE, warning= FALSE, message=FALSE}

scatter = final_data %>% 
  drop_na() %>% 
  ggplot(aes(x = year, y = mean_temp, color = region))+
  geom_point( alpha = 0.2)+
  geom_smooth( se = FALSE)+
  labs(title = "Average Temps(F) by Region from 1953-2018", 
       xlab = "Year",
       ylab = "Average Temperature(F)")+
  theme_bw()+
  theme(legend.position = "left")

ggplotly(scatter) %>% 
  layout(
    xaxis = list(title = "Year",
                 zeroline = TRUE),
    yaxis = list(title = "Avg Temp (F)",
                 zeroline = TRUE)
  )
```


This scatterplot showed the average temperature (F) from 1953 to 2018, by region. Overall, the southeast region had the highest average temperatures and the Northeast had the lowest average temperatures. Though it may be hard to see,
the average temperature increased from 70.81F to 72.28F from 1953 to 2018 in Florida, 59.25F to 62.26F in Arizona, 
56.63F to 55.23F in Missouri, and 53.54F to 54.14F in New Jersey.The different colored lines show the average trend over the years for average temperature, for each region.

## Sever Storm(s), Drought, Fire, and Flood by Region, 1953-2018

```{r, panel, echo = FALSE, warning= FALSE, message=FALSE}
panel = 
regional_disaster_count %>% 
  group_by(disaster, region)%>% 
  dplyr::select(region, disaster, year, count_region)%>% 
  rename(total = count_region) %>% 
  drop_na() %>% 
  filter(disaster == "Severe Storm(s)"| 
           disaster =="Drought"| 
           disaster =="Fire"|
           disaster =="Flood") %>% 
  ggplot(aes(x = year, y = total, color = disaster))+
  geom_line()+
  theme(aspect.ratio = 1, axis.text.x = element_text(angle = 90,  hjust = 1)) +
          facet_wrap(~region)
        
        
ggplotly(panel) %>% 
  layout(title = "Specific Disasters by Region",
         yaxis = list(title = "Disaster Count", 
                      zeroline = TRUE))
```

This panel shows the frequency of selected disaster types from 1953-2018 for each region. The disasters are shown in different colors. The West and Northeast regions had relatively few occurences of these specific disasters compared to the Midwest and Southeast regions. These four disaster types had the highest prevalence in our dataset which is why we focused on them.


## Total precipitation in inches for each region

```{r, echo = FALSE, warning= FALSE, message=FALSE}
final_data %>% 
  drop_na() %>% 
  dplyr::select(region, year, total_precip) %>% 
  distinct() %>% 
  mutate(text_label = str_c( year, "\nTotal Precip(in.):", total_precip)) %>% 
  plot_ly( x = ~year, y = ~total_precip, type = "scatter", mode = "markers",
    color = ~region, text = ~text_label, alpha = 0.5) %>% 
  layout(title = "Total Precipitation by Region, from 1953-2018",
         xaxis = list(title = "Year",
                      zeroline = TRUE),
         yaxis = list(title = "Total Precip (in.)",
                      zeroline = TRUE)) 

```

This plot shows the breakdown of the total precipitation amount in inches, by state for each year, and was split up into regions with the total precipitation summed up for each region. The Southeast had the most inches in precipitation overall, while the West had the least total amount of precipitation. 

## Total Number of Disasters by Region from 1953-2018, 

```{r, echo = FALSE, warning= FALSE, message=FALSE}
 final_data %>% 
  drop_na() %>% 
  dplyr::select( n_disaster_state, region, year) %>% 
  group_by(region) %>% 
  distinct() %>%
  mutate(
    text_label = str_c("total disasters: ", n_disaster_state, "\n", region)) %>%
  plot_ly(
    x = ~year, y = ~n_disaster_state, type = "bar", mode = "markers",
    color = ~region, text = ~text_label, alpha = 0.5) %>% 
  layout(title = "Total Disasters by Region, from 1953-2018",
         xaxis = list(title = "Year",
                      zeroline = TRUE),
         yaxis = list(title = "Total Number of Disastesr",
                      zeroline = TRUE))
```

This plot shows the number of disasters in each region from 1953 to 2018. We can see that the number of disasters has been on the rise and the total in the 2000s is larger compared to 1970-1980. In 2005, the Southeast had 763 disasters which is the spike seen in the graph. Overall, the Northeast and Western regions had the fewest number of disasters compared the Midwest and Southeast. 

## Number of Each Disaster by Region, 1953-2018

```{r bargraph, echo=FALSE, message=FALSE, warning=FALSE}
west = final_data %>% 
  filter(region == "west") %>% 
  group_by(disaster) %>% 
  drop_na() %>% 
  summarize(
    total_w = sum(n_disaster_state)
  ) 

midwest = final_data %>% 
  filter(region == "midwest") %>% 
  group_by(disaster) %>% 
  drop_na() %>% 
  summarize(
    total_mw = sum(n_disaster_state)
  ) 

southeast = final_data %>% 
  filter(region == "southeast") %>% 
  group_by(disaster) %>% 
  drop_na() %>% 
  summarize(
    total_se = sum(n_disaster_state)
  ) 

northeast = final_data %>% 
  filter(region == "northeast") %>% 
  group_by(disaster) %>% 
  drop_na() %>% 
  summarize(
    total_ne = sum(n_disaster_state)
  ) 

ne_se = dplyr::left_join(northeast, southeast, by = "disaster") 
mw_w = dplyr::left_join(midwest, west, by = "disaster")
us = dplyr::left_join(ne_se, mw_w, by = "disaster") %>% 
  rename(
    northeast = total_ne,
    southeast = total_se,
    midwest= total_mw,
    west = total_w
  ) %>% 
  pivot_longer(
    northeast:west,
    names_to = "region",
    values_to = "n_disaster"
  ) %>% 
  arrange(desc(n_disaster))

p = us %>% 
plot_ly( x = ~disaster, y = ~n_disaster, type = "bar",
         color = ~region) %>% 
  layout(
    title = "Number of Each Disasters by Region from 1953-2018",
         xaxis = list(title = "Type of Disaster",
                      zeroline = TRUE),
         yaxis = list(title = "Disaster Count",
                    zeroline = TRUE))
p
```

This bar graph shows the total number of different disasters for each region in the US, from 1953-2018. The individual number of disasters for each state are shown along the bars. We observed that the Southeast had the most hurricanes, while the midwest had a greater frequency of floods. Overall, it seems that the Midwest and Southeast have the higher frequency of disasters compared to the West and Northeast, which matches the trends we see in our other visualizations. 

 









